\documentclass{book}
\usepackage[left=2.5cm,top=3cm,right=2.5cm,bottom=3cm]{geometry}
\usepackage[colorlinks=true]{hyperref}
\usepackage{graphicx}
\usepackage{bsymb}
\usepackage{b2latex}
\usepackage{fancyhdr,lastpage,color}
\usepackage{verbatim}
\usepackage{wrapfig}
\usepackage{makeidx}

\input{config}
\input{makros}

\title{User Manual for Rodin v.\versionnr}
\author{
Work in Progress\\
Handbook $ $Rev$ $ \\
\\
\href{mailto:rodin-handbook@formalmind.com}{rodin-handbook@formalmind.com}\\
\href{http://handbook.event-b.org}{handbook.event-b.org}
}

\begin{document}        

\maketitle

\ifplastex
\else
\tableofcontents
\fi

\chapter{Introduction}

The Rodin documentation needs improvement.  This has been established at the Deploy Exec Meeting in 2010.  We've been collecting background information and the requirements in the \href{http://wiki.event-b.org/index.php/User_Documentation_Overhaul}{Rodin Wiki}.

What you see here is a working draft of the documentation.  We are grateful for any feedback that you may have (\ref{feedback}).

\section{Overview}

This handbook consists of three parts:

\begin{description}
	\item[Introduction] You are reading the introduction right now.  It helps you to orient yourself and to find information quickly.
	\item[Tutorial] If you are completely new to Rodin, the Tutorial is a good way to get up to speed quickly.  It guides you through installation and usage of the tool and gives you an overview of the Event-B modeling notation.
	\item[Reference] The reference section provides comprehensive documentation of Rodin, and its components.
	\item[Frequently Asked Questions] Common issues are listed by category in the FAQ.
	\item[Index] Particularly for the print version of this handbook, we included an index.  In the electronic versions, you may want to try the search functionality as well.
\end{description}

\subsection{Formats of this Handbook}

The handbook comes in various formats:

\begin{description}
	\item[Eclipse Help] Rodin is shipped with Rodin and can be accessed through the help system.  The handbook will be updated through the standard Rodin update mechanism.
	\item[Online Help] You can access the handbook online at \url{http://handbook.event-b.org}.
	\item[PDF Help] Both online versions also include a link to the PDF version of the handbook.
	\item[Phyiscal Book] If there is enough interest, we may make the handbook available through a print-on-demand retailer.
\end{description}

\subsection{Rodin Wiki}
\label{rodin_wiki}

This handbook is complemented by the Rodin wiki (\url{http://wiki.event-b.org/}).  Sometimes, the handbook will refer to the wiki for more information.  Also, plugin and developer information is usually located in the wiki.

\subsection{Feedback}
\label{feedback}

All online versions of the handbook contain a button or link for feedback.  Work on the handbook continues at least until January 2012, so your feedback will be read and will help to improve this handbook.  You can also submit feedback via email to \texttt{rodin-handbook@formalmind.com}.

\section{Foreword}

It would be nice to recruit somebody for the foreword - maybe Cliff and/or Jean-Raymond.

\section{Conventions}
\label{conventions}

We use the following conventions in this manual:

\tick{Checklists and Milestones are designated with tick. Here we summarize what we want to learn or should have learned so far.}
\info{Useful information and tricks are designated by the information sign.}
\warning{Potential problems and warnings are designated by a warning sign.}
\pencil{Examples and Code are designated by a pencil.}

We use \texttt{typewriter} font for file names and directories.

We use \textsf{sans serif font} for GUI elements like menus and buttons.  Menu actions are depicted by a chain of elements, separated by ``$\rangle$'', e.g. \textsf{File $\rangle$ New $\rangle$ Event-B Component}.

\section{Acknowledgements}
\label{sec:acknowledgements}

The content of this handbook has grown over many years, since the formation of the European Union IST Project RODIN in 2004.  Giving credit to every contributor is almost impossible, attempting to do so would almost certainly omit some people, which is not in the spirit of this work.  It should suffice to say that we extend our gratitude to all contributors to the Rodin Wiki (\ref{rodin_wiki}). In particular, we would like to thank Systerel\footnote{\url{http://www.systerel.fr}} for their significant contributions to the handbook, as they have been the main driver behind the tool and its documentation.

We would also like to thank Cliff Jones, who never gave up the quest for improved Rodin documentation.

The icons that you find throughout this handbook were created by Pixel-Mixer\footnote{\url{http://pixel-mixer.com/}}, who provides them freely.  Thanks!

\section{DEPLOY}
\label{deploy}

This work has been sponsored by the DEPLOY project\footnote{\url{http://www.deploy-project.eu/}}.  DEPLOY is an European Commission Information and Communication Technologies FP7 project.

The overall aim of the EC Information and Communication Technologies FP7 DEPLOY Project is to make major advances in engineering methods for dependable systems through the deployment of formal engineering methods. Formal engineering methods enable greater mastery of complexity than found in traditional software engineering processes. It is the central role played by mechanically-analysed formal models throughout the system development flow that enables mastery of complexity.

As well as leading to big improvements in system dependability, greater mastery of complexity also leads to greater productivity by reducing the expensive test-debug-rework cycle and by facilitating increased reuse of software.

The work of the project is driven by the tasks of achieving and evaluating industrial take-up, initially by DEPLOY's industrial partners, of DEPLOY's methods and tools, together with the necessary further research on methods and tools. 

\input{creative-commons}

\input{style-guide}

\input{tutorial}

\input{reference}

\input{faq}

\clearpage
\addcontentsline{toc}{chapter}{Index} 
\printindex

\end{document}

